On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}

### 2020-04-02 formal_glossary.tex
first full review (adding comments) of sec 3

The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
weakness is: ``Software weaknesses are errors that can lead to software
vulnerabilities. A software vulnerability, such as those enumerated on the
Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
can be directly used by a hacker to gain access to a system or network''.
The definition is circular if we interpret the word ``error'' and
``mistake'' with the same semantics: a weakness is an error that leads to a
vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
The only difference (between weakness and vulnerability) seems to be
that one can consider weakness as a ground term and state that a
vulnerability is caused by a weakness, i.e. 
$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
accepting (for now) the hierarchy in the CWE\autocite{CWE} as ground truth. 
Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability),
or the CVE reported in the NVD, as a ground truth.

\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
	A Region $\Weakness\subseteq\agent$ of an agent
	$\agent\in\agentuniverse$, is defined as Weakness\footnote{i.e.
	representing an error introduced by the agent into any phase of
	production, e.g.  of the secure process development life-cycle, of any
	system or subsystem} iff there exists $\world\in\possibleworlds$ such
	that $\world\models\Weakness$, $\world\models\Diamond\Vulnerability$,
	and $\neg\dr{\Weakness}{\Vulnerability}$; where $\Vulnerability$ is a
	Region of an agent $\agent\in\agentuniverse$.
\end{definition}

@@

\fixnote{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
\end{example}

@@

\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
	\emph{Defect} if the agent is referred to as Nature.
\end{definition}

@@

The formalization of the Exploitation and Black-hat Hacking processes is the
formalization of two foundational cybersecurity concepts which lay the basis
for the definition (and formalization) of the cybersecurity theory in the next
Section (Section~\ref{sec:theory}). The formal definition of those
processes shall:
\begin{itemize}
	\item in Section~\ref{sec:theory}, formally define a system as a structure of agents, and
	\item in Section~\ref{sec:engineering}, correlate the formal definition of agents (i.e. the
		Definition~\ref{def:agent}) with the Causality principle (i.e.
		the formalization of the Kripke structure in
		Definition~\ref{def:modallogic}). 
\end{itemize}

>>>>

The definition given by the MITRE in\autocite{MITRE2020CWEweakness} of
weakness is: ``Software weaknesses are errors that can lead to software
vulnerabilities. A software vulnerability, such as those enumerated on the
Common Vulnerabilities and Exposures (CVE) List, is a mistake in software that
can be directly used by a hacker to gain access to a system or network''.
The definition is circular if we interpret the word ``error'' and
``mistake'' with the same semantics: a weakness is an error that leads to a
vulnerability and a vulnerability is a mistake which, in turn, is a weakness.
The only difference (between weakness and vulnerability) seems to be
that one can consider weakness as a ground term and state that a
vulnerability is caused by a weakness, i.e. 
$\World,\Weakness\models\Diamond\Vulnerability\wedge\Weakness$ where
$\Weakness,\Vulnerability$ are Regions of Weaknesses and Vulnerabilities (resp.);
accepting (for now) the hierarchy in the CWE\autocite{CWE} as ground truth. 
Similarly, we consider the CVE\autocite{CVE} (a database of Vulnerability),
or the CVE reported in the NVD, as a ground truth. 
\fix{mr}{it needs to be stressed that without the formalization of
the concept of error, we cannot properly formalize the concept of weakness
which remains just a region as any other region with the only 
constraint that it must be connected to another region called Vulnerability (which 
in turn depend on the definition of error)}

\fix{mr}{the following definition should be moved (and detailed)
to section~\ref{engineering} when the concept of weakness is defined}
\begin{definition}{\bf Region: Weakness and Vulnerability --}\label{def:weakness}
	A Region $\Weakness\subseteq\agent$ of an agent
	$\agent\in\agentuniverse$, is defined as Weakness\footnote{i.e.
	representing an error introduced by the agent into any phase of
	production, e.g.  of the secure process development life-cycle, of any
	system or subsystem} iff there exists $\world\in\possibleworlds$ such
	that $\world\models\Weakness$, $\world\models\Diamond\Vulnerability$,
	and $\neg\dr{\Weakness}{\Vulnerability}$; where $\Vulnerability$ is a
	Region of an agent $\agent\in\agentuniverse$.
\end{definition}

@@

\fix{mr}{shouldn't we consider the agent who introduces the weakness as dishonest? don't we say this in part 1?}
\end{example}

@@

\fix{mr}{similar to weakness and vuln, the following definition should be kept here just to show that
it is impossible to defined those concept without the proper definition of the concept of error. and the definition should
be introduced in section~\ref{engineering} when the concept of weakness is defined}
\begin{definition}{\bf Region: Vulnerability and Defect --}\label{def:vulnerability-defect}
A region $\region\subseteq\agent$ of an agent $\agent\in\agentuniverse$ is
	called \emph{Vulnerability} if the agent $\agent$ is referred to as Mankind, 
	\emph{Defect} if the agent is referred to as Nature.
\end{definition}

@@

The formalization of the Exploitation and Black-hat Hacking processes is the
formalization of two foundational cybersecurity concepts which lay the basis
for the definition (and formalization) of the cybersecurity theory in the next
Section (Section~\ref{sec:theory}). The formal definition of those
processes shall:
\begin{itemize}
	\item in Section~\ref{sec:theory}, formally define a system as a structure of agents, and
	\item in Section~\ref{sec:engineering}, correlate the formal definition of agents (i.e. the
		Definition~\ref{def:mankind-nature}) with the Causality principle (i.e.
		the formalization of the Kripke structure in
		Definition~\ref{def:modallogic}). 
\end{itemize}
